---
title: "Caregiver Safety Trends: Advanced Dashboard"
author: "Harlie Lockhart"
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
    vertical_layout: fill
runtime: shiny
---

```{r setup, include=FALSE}
library(flexdashboard)
library(tidyverse)
library(readxl)
library(plotly)
library(shiny)
library(DT)

knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)


Caregiver_Injury_Dataset <- read_excel("Caregiver Injury Dataset.xlsx")

Caregiver_Injury_Dataset <- Caregiver_Injury_Dataset %>%
  mutate(
    Severity = case_when(
      Incident_Type %in% c("Needlestick", "Physical Assault") ~ "Severe",
      Incident_Type %in% c("Lifting Injury", "Slip/Fall") ~ "Moderate",
      TRUE ~ "Low"
    )
  )


DT::datatable(Caregiver_Injury_Dataset)

getwd()

```


# Overview

## Caregiver Safety Trends Dashboard

### Purpose
This dashboard explores **simulated caregiver safety incident data** to identify patterns across:
- Hospital units
- Work shifts
- Incident types

The goal is to support **proactive safety planning**, staffing awareness, and training prioritization.

---

### How to Use This Dashboard
- Navigate to the **Interactive Insights** tab to filter incidents by unit and shift.
- KPIs and visualizations will update dynamically based on your selections.
- Use the insights to identify high-risk areas and trends.

---

### Ethical & Data Disclaimer
- All data shown is **simulated** and used for educational purposes only.
- No real patient or caregiver information is included.
- Trends are illustrative, not diagnostic.

---

### Author
**Harlie Lockhart**  
University of Arkansas
Data Science (Biomedical & Healthcare Informatics)
Minor in Medical Humanities

---

### Contact & Feedback
Questions, feedback, or recommendations related to caregiver safety and dashboard design are welcome.

📧 **Email:** [harliel@uark.edu](mailto:harliel@uark.edu)

---

### Use & Copyright Notice
© 2026 **Harlie Lockhart**. All rights reserved.

This dashboard and its contents are intended for **educational use only**.  
Reproduction, redistribution, or reuse of this material without permission is not allowed.


# Interactive Insights

```{r}

### Filters
selectInput("unit_filter", "Select Unit:", choices = c("All", unique(Caregiver_Injury_Dataset$Unit)), selected = "All")
selectInput("shift_filter", "Select Shift:", choices = c("All", unique(Caregiver_Injury_Dataset$Shift)), selected = "All")

### Summary Metrics
renderUI({
  data <- filtered_data()
  tagList(
    div(strong("Total Incidents:"), nrow(data)),
    div(strong("Unique Units:"), n_distinct(data$Unit)),
    div(strong("Unique Shifts:"), n_distinct(data$Shift)),
    div(strong("Incident Types:"), n_distinct(data$Incident_Type)),
    div(strong("Severity Levels:"), n_distinct(data$Severity))
  )
})

### Filtered Dataset Table
renderTable({
  filtered_data()
})

# --- Reactive filtered dataset 
filtered_data <- reactive({
  data <- Caregiver_Injury_Dataset
  if (input$unit_filter != "All") {
    data <- data %>% filter(Unit == input$unit_filter)
  }
  if (input$shift_filter != "All") {
    data <- data %>% filter(Shift == input$shift_filter)
  }
  data
})

```

# Key Performance Indicators (KPI's)

```{r}
renderUI({
data <- filtered_data()
total_incidents <- nrow(data)
total_units <- n_distinct(data$Unit)
total_shifts <- n_distinct(data$Shift)
total_types <- n_distinct(data$Incident_Type)
total_severity <- n_distinct(data$Severity)

tagList(
div(strong("Total Incidents: "), total_incidents),
div(strong("Units Reported: "), total_units),
div(strong("Shifts Reported: "), total_shifts),
div(strong("Incident Types: "), total_types),
div(strong("Severity Levels: "), total_severity)
)
})
```

```{r}
# Incidents by Unit
renderPlotly({
data <- filtered_data() %>% count(Unit)
p <- ggplot(data, aes(x = Unit, y = n, fill = ifelse(n == max(n), "red", "gray70"))) +
geom_col(width = 0.5) +
labs(title = "Caregiver Safety Incidents by Unit", x = "Unit", y = "Number of Incidents") +
scale_fill_identity() +
theme_minimal()
ggplotly(p)
})
```

# Incidents by Shift

```{r}
renderPlotly({
data <- filtered_data() %>% count(Shift)
p <- ggplot(data, aes(x = Shift, y = n, fill = "gray70")) +
geom_col(width = 0.5) +
labs(title = "Caregiver Safety Incidents by Shift", x = "Shift", y = "Number of Incidents") +
theme_minimal()
ggplotly(p)
ggplotly(p, height = 500) %>%
    layout(yaxis = list(autorange = TRUE),
           margin = list(l = 150, r = 100, t = 100, b = 100))
})

```

# Most Common Incident Types

```{r}
renderPlotly({
  data <- filtered_data() %>% count(Incident_Type) %>% arrange(desc(n))
  p <- ggplot(data, aes(x = reorder(Incident_Type, n), y = n,
                        fill = ifelse(n == max(n), "red", "gray70"))) +
    geom_col(width = 0.5) +
    coord_flip() +
    scale_y_continuous(expand = expansion(mult = c(0.05, 0.2))) +
    labs(title = "Most Common Caregiver Safety Incidents", 
         x = "Incident Type", y = "Number of Incidents") +
    scale_fill_identity() +
    theme_minimal()
  
  ggplotly(p, height = 400) %>%
    layout(yaxis = list(autorange = TRUE),
           margin = list(l = 100, r = 50, t = 50, b = 50))
})


```


# Severity Analysis 

```{r}
selectInput(
  "severity_filter",
  "Select Severity Level:",
  choices = c("All", unique(Caregiver_Injury_Dataset$Severity)),
  selected = "All"
)

severity_data <- reactive({
  data <- Caregiver_Injury_Dataset
  if (input$severity_filter != "All") {
    data <- data %>% filter(Severity == input$severity_filter)
  }
  data
})

renderPlotly({
  data <- severity_data() %>% count(Unit, Severity)
  p <- ggplot(data, aes(x = Unit, y = n, fill = Severity)) +
    geom_col() +
    labs(
      title = "Incidents by Unit and Severity",
      x = "Unit",
      y = "Number of Incidents"
    ) +
    theme_minimal()
  ggplotly(p, height = 400) %>%
    layout(yaxis = list(autorange = TRUE),
           margin = list(l = 100, r = 50, t = 50, b = 50))
})
  
```

# Key Insights & Recommendations

### Key Findings

-   Med-Surg units reported the highest number of incidents.
-   Night shifts had slightly more incidents, but day shifts were still high.
-   Needlesticks were the most common incident type; verbal aggression and lifting injuries tied for second.

### Operational Recommendations

-   Targeted reminders and training for high-incident units and shifts.
-   Staffing adjustments and workload review for night shifts.
-   Consistent reporting supports proactive safety planning.

### Ethical Disclaimer

All data is **simulated** for illustrative purposes; no real patient or caregiver data was used.




